---
title: "TSA - RBOB"
author: "Bo Sang, Danial Sheikh, Eduardo Vega"
date: "`r Sys.Date()`"
output: html_document
resource_files:
- TSA_template.Rmd
- TSA_template.Rmd
---

<style type="text/css"> body, td {font-size: 12px;} code.r{font-size: 10px;} pre {font-size: 10px} </style>

```{r, echo = F}
knitr::opts_chunk$set(echo = T,warning = F,message = F, fig.width = 4, fig.height = 3,fig.align = "center",tidy = FALSE, strip.white = TRUE)
```

**this template is a guide to organize your project in a clear manner**

**quality of presentation refers amongst others to showing code, tables and charts in a manner that flows with your story. Details and computations shall remain in the code chunks without being rendered into the html output.**

## Experimentation Part

+ Do all your practice and experimentation prior to starting the assignment in this section.

**Disregard the sections below for the time being**

## Summary

### What's changed

### Current Exposure

### New Trades if any

## Form a Market View

### Supply Demand Indicators

Build a SD balance dashboard to support forming and explaining a market view below.

```{r}
# You are expected to build your own SD balance from the EIA and other website APIs
# The following is an example of series names for Crude SD components to get you going.
# It is by no means a substitute for doing your own research
library(RTL)
data("tickers_eia")
tickers_eia %>% dplyr::filter(product == "crude")
```

+ Clearly laid out components of SD balances.
+ State why they matter.
+ Provide both long term and short term views. Your trade horizon is weekly - can you identify clearly from the chart when a change occurs?

### Market View

+ View based on SD indicators ensuring that your conclusions are clearly readable from your charts.
+ Build analytics that is relevant to your current view.
+ If you have explored more, put them in an appendix. 

## Desired Exposure

Translating your market view into strength of market call and allocate risk versus maximum limit

+ Evidence of converting market view into a strength of market call.
+ Strength of market call translates into allocating risk allocation vs your limits.

## Monetization Strategies

+ Evidence of exploring what we have covered in class.
+ How do you take those further with exploring various combinations along the grade, delivery location and delivery timing axis.
+ Diligence in being nimble in those in light of evolving market context.

## Risk Appetite

+ Allocate risk in the context of risk reward ensuring capital preservation.
+ Evidence of a basic framework that you utilize to make your decision. Best is if you can articulate it clearly and quantify it.

## Execution

+ What trades are you executing? 
+ Rationale stated clearly?
+ Entry/ Exit levels stated?
+ 

## Profit and Loss ("PL" Attribution

For each strategy:

+ Provide a grid of PL attribution by risk factors (flat price, time spread, crack,) 
+ Is it in line with your market call? Learnings...

## Lessons Learned

What have you learned from this project?

## Questions for Weekly Meeting with Prof

1. a...
2. b...






